---
title: "Adventures in Bayesian Structural Time Series"
author: "Andrew Bates, Josh Gloyd, Tyler Tucker"
urlcolor: blue
output: beamer_presentation
theme: 'dnd'
#colortheme: 'dndcolors'
#fonttheme: 'dndfonts'
---

## What are Time Series?
```{r, echo = FALSE, message = FALSE, fig.align = "left", fig.width = 8, fig.height = 4, out.width = "55%"}
library(readr)
gilbralter <- read_csv("data/gilbraltersimple.csv")  
gilb <- ts(gilbralter$tempMean, start=c(2004,1,13), end=c(2017, 11, 25), frequency=30)
plot(gilb, main = 'Gibralter Sea Surface Temperature', ylab = 'SST')
```

```{r, echo = FALSE, message = FALSE, fig.align = "right", fig.width = 4, fig.height = 4, out.width = "35%"}
library(knitr)
img1_path <- "images/nowcasting.png"
include_graphics(img1_path)
```

## What are BSTS?

> - [Predicting the Present with Bayesian Structural Time Series](http://www.inderscience.com/offer.php?id=59942) by Steven L. Scott and Hal Varian (Google)

> - Implementation
>      - R: [bsts](https://cran.r-project.org/web/packages/bsts/index.html)
>          - or [Causal Impact](https://google.github.io/CausalImpact/CausalImpact.html)
>      - Python: [Causal Impact](https://github.com/tcassou/causal_impact)

> - Comprised of 3 components:
>      - Structural Time Series model (a.k.a. state space model)
>      - Spike and Slab regression
>      - Bayesian model averaging

## Structural Time Series 

- Data from unobserved **state space** plus noise
- Model the latent state space instead of the data directly

### Local Level Model

- $y_t$: data
- $\mu_t$: latent state

\begin{align*}
y_t &= \mu_t + \varepsilon_t  &\varepsilon_t \sim N(0,\sigma^2_{\varepsilon}) \\
\mu_{t+1} &= \mu_t + \xi_t  &\xi_t \sim N(0, \sigma^2_{\xi}) 
\end{align*}

Analogous to the inercept in linear regression but allowing for the intercept to vary over time


## Structural Time Series

### Local Linear Trend Model

- $y_t, \mu_t$: same as before
- $\nu_t$: slope (additional state component)

\begin{align*}
y_t &= \mu_t + \varepsilon_t  &\varepsilon_t \sim N(0, \sigma_{\varepsilon}^2) \\
\mu_{t+1} &= \mu_t + \nu_t + \xi_t  &\xi_t \sim N(0, \sigma_{\xi}^2) \\
\nu_{t+1} &= \nu_t + \zeta_t  &\zeta_t \sim N(0, \sigma_{\zeta}^2)
\end{align*}

## Structural Time Series

### General Form

- $y_t$: data
- $\alpha_t$: state component

\begin{align}
y_t &= Z_t'\alpha_t + \varepsilon_t  &\varepsilon_t \sim N(0, H_t)  \\
\alpha_{t+1} &= T_t\alpha_t + R_t\eta_t  &\eta_t \sim N(0, Q_t) 
\end{align}

- (1): **observation equation**
- (2): **transition equation**


## Bayesian Context

- Spike and slab regression
    - Used when regression components are included
    - Variable selection technique
    - Prior on regression coefficients

- Bayesian Model Averaging
    - Consequence of spike and slab prior
    - Different $\beta$s included in each draw of posterior (i.e. different model on each draw)

- Prior Elicitation and Posterior Sampling
    - Inclusion probabilities for regression coefficients 
    - Or: expected model size, expected $R^2$, weight given to $R^2$
    - Gibbs sampler (stochastic search variable selection) to draw from posterior
    - For details see paper by Scott and Varian

## Spike and Slab Regression

### Joint Prior

- $\gamma_k = 1$ if $\beta_k \neq 0$
- $\gamma_k = 0$ if $\beta_k = 0$

\begin{align*}
p(\beta,\gamma,\sigma^2_{\epsilon}) = p(\beta_{\gamma}|\gamma,\sigma^2_{\epsilon})p(\sigma^2_{\epsilon}|\gamma)p(\gamma)
\end{align*}

## Spike and Slab Regression

### "Spike" Prior

- $\pi_k$: probability a particular $\beta_k$ is included in the model

\begin{align*}
\gamma \sim \prod_{k=1}^{K} \pi_k^{\gamma_k}(1-\pi_k)^{1-\gamma_k}
\end{align*}

## Spike and Slab Regression

### "Slab" Prior

- $b_{\gamma}$: b is a vector of prior means
- $\Omega_{\gamma}^{-1}$: denotes the rows and columns of symmetric matrix $\Omega^{-1}$ corresponding to $\gamma_k = 1$
- $ss$: prior sum of squares
- $\nu$: prior sample size

\begin{align*}
\beta_{\gamma}|\sigma^2_{\epsilon},\gamma \sim N\Big(b_{\gamma},\sigma^2_{\epsilon}(\Omega_{\gamma}^{-1})^{-1}\Big) \hspace{10mm} \frac{1}{\sigma^2_{\epsilon}}|\gamma \sim Ga\Big(\frac{\nu}{2},\frac{ss}{2}\Big)
\end{align*}

