---
title: "Implementing BSTS Forecasting"
author: "Andrew Bates, Josh Gloyd, Tyler Tucker"
urlcolor: blue
output: beamer_presentation
theme: '/../dnd'
#colortheme: 'dndcolors'
#fonttheme: 'dndfonts'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "normal")
library(Boom)
library(zoo)
library(bsts)     # load the bsts package
library(readr)
library(latex2exp)
options( warn = -1 )
options(repr.plot.width=7, repr.plot.height=5)
```

## Time Series Visual

```{r, echo=TRUE, results="hide", message = FALSE, fig.align = "left", fig.width = 8, fig.height = 4, out.width = "75%"}
gilbralter <- read_csv("data/gilbralter.csv")
gilT <- ts(gilbralter$tempMean, start=c(2004,1,13),
           end=c(2017, 11, 25), frequency=30)
plot(gilT, main='SST of Gilbralter region',
    xlab='Date',
    ylab='Temperature [C]')
```

## Set up BSTS model?
These three lines of code are setting up the model as 
$$
y_t = \mu_t + \tau_t + \epsilon_t
$$
where $\mu_t$ is a local linear component, $\tau_t$ is a seasonal component, and $\epsilon_t$ is a white noise process centered about zero and with standard deviation $\sigma^2$. 
```{r, echo=TRUE, results="hide", message = FALSE, fig.align = "left", fig.width = 6, fig.height = 4, out.width = "65%"}
nseasons = 30
ss <- list()
ss <- AddLocalLinearTrend(ss, gilT)
ss <- AddSeasonal(ss, gilT, nseasons=nseasons)
```
## Run the model
Implement the bsts function with time series, state specification object, and number of iterations runs the MCMC and fits the parameters.
```{r, echo=TRUE, results="hide", message = FALSE, fig.align = "left", fig.width = 8, fig.height = 4, out.width = "75%"}
modelGilT <- bsts(gilT,
               state.specification=ss,
               niter = 1000)
```
## Interpret
```{r, echo=TRUE, eval=FALSE}
plot(modelGilT,
    main=TeX('Conditional Expectation of $y_i$'),
    xlab='time idx', ylab='distribution', 
    ylim=c(15, 26))
legend("topright",
       legend=c(TeX('$E(y_i|data)$'), "data"),
       col=c("black", "blue"),
       lty=c(1,NA), pch=c(NA, 1), cex=.7)
```
## Interpret
```{r, echo=FALSE, results="hide", message = FALSE, fig.align = "left", fig.width = 8, fig.height = 4, out.width = "95%"}
plot(modelGilT,
    main=TeX('Conditional Expectation of $y_i$'),
    xlab='time idx', ylab='distribution', 
    ylim=c(15, 26))
legend("topright",
       legend=c(TeX('$E(y_i|data)$'), "data"),
       col=c("black", "blue"),
       lty=c(1,NA), pch=c(NA, 1), cex=.7)
```
## Interpret
Trend and Season plots are available too
```{r, echo=TRUE, results="show", message = FALSE, fig.align = "left", fig.width = 8, fig.height = 4, out.width = "75%"}
plot(modelGilT, "components")
```
## Forecasting
From here we can perform forecasting 12 time steps in the future (144 days) with the following code.
Green dotted line is the 97.5 % prediction interval, the blue line is the 2.5 % PI
```{r, echo=TRUE, results="show", message = FALSE, fig.align = "left", fig.width = 8, fig.height = 4, out.width = "75%"}
pred1 <- predict(modelGilT, horizon = 12)
plot(pred1, plot.original = 156)
```