---
title: "Adventures in Bayesian Structural Time Series"
subtitle: "Part 3: Analyzing SST Data"
author: "Andrew Bates, Josh Gloyd, Tyler Tucker"
urlcolor: blue
header-includes:
  - \usepackage{themes/beamerthemednd}
output: 
  beamer_presentation:
    incremental: false
#theme: 'dnd'
#colortheme: 'dndcolors'
#fonttheme: 'dndfonts'
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(fig.align = 'center', out.width = '90%',
                      message = FALSE, cache = TRUE)
```


<!-- slide 3 -->

## Outline

- SST data.
- Model fitting of SST data, whose components include: local level, local linear trend, and seasonal trend.
- Generating a posterior distribution of our model time series
- Some heuristic comparisons between the different models.

<!-- slide 4 -->

## SST Data

```{r, echo = FALSE}
library(readr)
library(bsts) # also loads BoomSpikeSlab, Boom, MASS, zoo, xts


gilbralter <- read_csv("data/gilbraltersimple.csv")
gilt <- ts(gilbralter$tempMean, start=c(2004,1,13),
           end=c(2017, 11, 25), frequency=30)
plot(gilt, main='SST of Gilbralter region',
     xlab='Date',
     ylab='Temperature (C)')
```

<!-- slide 5 -->
    
## SST Data

- SST data come from Argo floats
- Aggregated every 12 days
- January 2004 to November 2017
- Obtained from [www.Argovis.com](http://argovis.com/map)
- Learn more about Argovis by watching a tutorial at [https://www.youtube.com/watch?v=IlNJ0owuTHM&t=0s](https://www.youtube.com/watch?v=IlNJ0owuTHM&t=0s)
<!-- slide 6 -->

## Setup

```{r, eval = FALSE}
library(readr)
library(bsts) 
# bsts also loads BoomSpikeSlab, Boom, MASS, zoo, xts


gilbralter <- read_csv("data/gilbraltersimple.csv")
gilt <- ts(gilbralter$tempMean, start=c(2004,1,13),
           end=c(2017, 11, 25), frequency=30)
plot(gilt, main='SST of Gilbralter region',
     xlab='Date',
     ylab='Temperature (C)')
```


<!-- slide 7 -->

## Model Fitting

### Local Level Model

\begin{align*}
y_t &= \mu_t + \varepsilon_t  &\varepsilon_t \sim N(0,\sigma^2_{\varepsilon}) \\
\mu_{t+1} &= \mu_t + \xi_t  &\xi_t \sim N(0, \sigma^2_{\xi}) 
\end{align*}


###

```{r, eval = FALSE}
ll_ss <- list()
ll_ss <- AddLocalLevel(state.specification = ll_ss,
                       y = gilt)
ll_fit <- bsts(gilt, state.specification = ll_ss,
               niter = 1e3)
```


<!-- slide 8 -->

## Model Plotting

```{r, echo = FALSE}
ll_ss <- list()
ll_ss <- AddLocalLevel(state.specification = ll_ss,
                       y = gilt)
ll_fit <- bsts(gilt, state.specification = ll_ss,
               niter = 1e3, ping = 0)
```


```{r}
plot(ll_fit)
```

<!-- slide 9 -->

## Model Plotting

```{r}
plot(ll_fit, 'components')
```


<!-- slide 10 -->

## Model Plotting

```{r}
plot(ll_fit, 'residuals')
```


<!-- slide 11 -->

## Forecasting

```{r}
ll_pred <- predict(ll_fit, horizon = 30)  
plot(ll_pred, plot.original = 90)  
```

<!-- slide 12 -->

## Model Fitting

### Local Linear Trend Model

\begin{align*}
y_t &= \mu_t + \varepsilon_t  &\varepsilon_t \sim N(0, \sigma_{\varepsilon}^2) \\
\mu_{t+1} &= \mu_t + \nu_t + \xi_t  &\xi_t \sim N(0, \sigma_{\xi}^2) \\
\nu_{t+1} &= \nu_t + \zeta_t  &\zeta_t \sim N(0, \sigma_{\zeta}^2)
\end{align*}


###

```{r, eval = FALSE}
llt_ss <- list()
llt_ss <- AddLocalLinearTrend(
  state.specification = llt_ss, y = gilt)
llt_fit <- bsts(gilt, state.specification = llt_ss,
                niter = 1e3)
```


<!-- slide 13 -->

## Forecasts

```{r, echo = FALSE}
llt_ss <- list()
llt_ss <- AddLocalLinearTrend(
  state.specification = llt_ss, y = gilt)
llt_fit <- bsts(gilt, state.specification = llt_ss,
                niter = 1e3, ping = 0)
```

```{r}
llt_pred <- predict(llt_fit, horizon = 30)
plot(llt_pred, plot.original = 90)
```


<!-- slide 14 -->

## Model Fitting

### Local Trend With Seasonality

\begin{align*}
&y_t = \mu_t + \tau_t + \varepsilon_t  &\zeta_t \sim N(0, \sigma^2_{\varepsilon}) \\
&\tau_t = - \sum_{s=1}^{S-1} \tau_{t-s} + \omega_t  &\tau_t \sim N(0, \sigma^2_{\omega})
\end{align*}

### 
```{r, eval = FALSE}
lts_ss <- list()
lts_ss <- AddLocalLinearTrend(lts_ss, y = gilt)
lts_ss <- AddSeasonal(lts_ss, gilt, nseasons = 30)
lts_fit <- bsts(gilt, state.specification = lts_ss,
                niter = 1e3)
```


<!-- slide 15 -->

## Components

```{r, echo = FALSE}
lts_ss <- list()
lts_ss <- AddLocalLinearTrend(lts_ss, y = gilt)
lts_ss <- AddSeasonal(lts_ss, gilt, nseasons = 30)
lts_fit <- bsts(gilt, state.specification = lts_ss,
                niter = 1e3, ping = 0)
```

```{r}
plot(lts_fit, 'components')
```


<!-- slide 16 -->

## Forecasts

```{r}
lts_pred <- predict(lts_fit, horizon = 30)
plot(lts_pred, plot.original = 90)
```



<!-- slide 17 -->

## Model Comparison

```{r}
CompareBstsModels(model.list = list(
  level = ll_fit, trend = llt_fit, season = lts_fit))
```

