---
title: "Adventures in Bayesian Structural Time Series"
subtitle: "Part 1: Introduction"
author: "Andrew Bates, Josh Gloyd, Tyler Tucker"
urlcolor: blue
header-includes:
  - \usepackage{themes/beamerthemednd}
output: 
  beamer_presentation:
#theme: 'dnd'
#colortheme: 'dndcolors'
#fonttheme: 'dndfonts'
---


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(fig.align = 'center', out.width = '95%')
```




<!-- slide 3 -->

## The Adventure

> - Structural time series
> - Bayesian approach to structural time series
> - Implementation via `bsts` in `R`


<!-- slide 4 -->

## Useful Resources

> - "Predicting the Present with Bayesian Structural Time Series"
Steven L. Scott and Hal Varian

> - "An Introduction to State Space Time Series Analysis"
Commandeur and Koopman

> - "Time Series Analysis By State Space Methods"
Durbin and Koopman

> - `bsts` documentation

> - Adventures in BSTS GitHub


<!-- slide 5 -->

## Time Series Review


> - Stochastic process indexed by time 
>     $\{X_t, \hspace{3pt} t \in \mathbb{T}\}$
    
> - (weak) Stationarity

$$E[X_t] = \mu$$

$$Cov(X_t, X_{t+k}) = \gamma(k)$$
    
> - Not i.i.d.


<!-- slide 6 -->

## ARMA Models

### AR(1)

\[X_t = \phi X_{t-1} + e_t\]
$e_t$ i.i.d. $(0,\sigma^2)$
$$|\phi| < 1$$


<!-- slide 7 -->

## AR(1)

```{r, echo = FALSE}
x <- arima.sim(model = list(ar = 0.42), n = 100, sd = 1)
plot(x, main = 'Simulated AR(1)')
```


<!-- slide 8 -->

## ARMA Models

### MA(1)

\[X_t = \theta e_{t-1} + e_t \]
$e_t$ i.i.d. $(0, \sigma^2)$


<!-- slide 9 -->

## MA(1)

```{r, echo = FALSE}
x <- arima.sim(model = list(ma = 0.42), n = 100, sd = 1)
plot(x, main = 'Simulated MA(1)')
```


<!-- slide 10 -->

## ARMA Models

### ARMA(p,q)

\[\phi(B)X_t = \theta(B)e_t\]

$$\phi(B)X_t = X_t - \phi_1X_{t-1} - \cdots - \phi_pX_{t-p}$$

$$\theta(B)e_t = e_t + \theta_1e_{t-1} + \cdots + \theta_qe_{t-q}$$


<!-- slide 11 -->

##

**Your quest...**



