---
title: "Adventures in Bayesian Structural Time Series"
subtitle: "Part 2: Structural Time Series"
author: "Andrew Bates, Josh Gloyd, Tyler Tucker"
urlcolor: blue
header-includes:
  - \usepackage{themes/beamerthemednd}
output: 
  beamer_presentation:
#theme: 'dnd'
#colortheme: 'dndcolors'
#fonttheme: 'dndfonts'
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center', out.width = '95%',
                      message = FALSE)
```


## Outline

> - Structural Time Series Models:
>     - Local level model
>     - Local linear trend model
>     - Models with a seasonal component
>     - Models with a regression component


<!-- slide 3 -->

## Structural Time Series 

> - Also called State Space Models
> - Data comes from unobserved variable called the  **state space**
> - We model the state space instead of the observed data directly


<!-- slide 4 -->

## Structural Time Series

### Local Level Model

- $y_t$: observed data
- $\mu_t$: unobserved state

- \begin{align*}
y_t &= \mu_t + \varepsilon_t  &\varepsilon_t \sim N(0,\sigma^2_{\varepsilon}) \\
\mu_{t+1} &= \mu_t + \xi_t  &\xi_t \sim N(0, \sigma^2_{\xi}) 
\end{align*}


<!-- slide 5 -->

## Local Level Model

```{r}
ll_sim <- function(n = 100, sigma = c(10, 5)){
  sig_eps <- sigma[1]
  sig_xi <- sigma[2]
  y <- numeric(n)
  mu <- numeric(n)
  mu[1] <- rnorm(1, sd = sig_xi)
  for(t in 1:n){
    mu[t+1] <- mu[t] + rnorm(1, sd = sig_xi)
    y[t] <- mu[t] + rnorm(1, sd = sig_eps)
  }
  return(ts(y))
}

y <- ll_sim()
plot(y, main = 'Simulated Local Level Model')
```



<!-- slide 6 -->

## Structural Time Series

### Local Linear Trend Model

- $y_t, \mu_t$: same as before
- $\nu_t$: slope (additional state component)

- \begin{align*}
y_t &= \mu_t + \varepsilon_t  &\varepsilon_t \sim N(0, \sigma_{\varepsilon}^2) \\
\mu_{t+1} &= \mu_t + \nu_t + \xi_t  &\xi_t \sim N(0, \sigma_{\xi}^2) \\
\nu_{t+1} &= \nu_t + \zeta_t  &\zeta_t \sim N(0, \sigma_{\zeta}^2)
\end{align*}


<!-- slide 7 -->

## Local Linear Trend Model

```{r}
set.seed(42)
llt_sim <- function(n = 100, sigma = c(10, 10, 10)){
  sig_eps <- sigma[1]
  sig_xi <- sigma[2]
  sig_zeta <- sigma[3]
  y <- numeric(n)
  mu <- numeric(n)
  nu <- numeric(n)
  mu[1] <- rnorm(1, sd = sig_xi)
  nu[1] <- rnorm(1, sd = sig_zeta)
  for(t in 1:n){
    nu[t+1] <- nu[t] + rnorm(1, sd = sig_zeta)  # trend
    mu[t+1] <- mu[t] + nu[t] + rnorm(1, sd = sig_xi)  # level
    y[t] <- mu[t] + rnorm(1, sd = sig_eps)
  }
  return(ts(y))
}

y <- llt_sim()
plot(y, main = 'Simulated Local Linear Trend Model')
```



<!-- slide 8 -->

## Structural Time Series

### Local Trend With Seasonality

- $\mu_t$: local linear trend
- $\tau_t$: seasonal component
    - $S$ dummy variables (1 for each season)

- \begin{align*}
&y_t = \mu_t + \tau_t + \varepsilon_t  &\zeta_t \sim N(0, \sigma^2_{\varepsilon}) \\
&\tau_t = - \sum_{s=1}^{S-1} \tau_{t-s} + \omega_t  &\tau_t \sim N(0, \sigma^2_{\omega})
\end{align*}


<!-- slide 9 -->

## Local Trend With Seasonality

```{r}
library(readr)
gilbralter <- read_csv("data/gilbraltersimple.csv")  
gilb <- ts(gilbralter$tempMean, start=c(2004,1,13), end=c(2017, 11, 25), frequency=30)
plot(gilb, main = 'Gibralter Sea Surface Temperature', ylab = 'SST')
```


<!-- slide 10 -->

## Structural Time Series

### Local Trend With Seasonality and Regression

- $\mu_t$: local linear trend 
- $\tau_t$: seasonal component
- $\beta_t^Tx_t$: regression component

- \begin{align*}
&y_t = \mu_t + \tau_t + \beta_t^Tx_t  + \varepsilon_t &\varepsilon_t \sim N(0, \sigma^2_{\varepsilon})
\end{align*}


<!-- slide 11 -->


## Structural Time Series

- (1): **observation equation**
- (2): **transition equation**


### General Form

- $y_t$: data
- $\alpha_t$: state variable

- \begin{align}
y_t &= Z_t'\alpha_t + \varepsilon_t  &\varepsilon_t \sim N(0, H_t)  \\
\alpha_{t+1} &= T_t\alpha_t + R_t\eta_t  &\eta_t \sim N(0, Q_t) 
\end{align}





